Many personal-agent systems now include memory, and many self-evolving agent
systems treat reusable skill acquisition as the primary path to improvement.
For personal AI, both are downstream of a more central problem: how to maintain
correctable understanding of one person over time. Elephant Agent is a
\emph{personal-model-first} memory architecture for personal AI. It centers a
Personal Model: a durable, grounded understanding across four lenses --- who
the user is (\textit{Identity}), what surrounds them (\textit{World}), what is
alive right now (\textit{Pulse}), and what their path has taught them
(\textit{Journey}).

The system treats memory as five cooperating layers rather than a transcript
archive or a vector store: Matriarch Core / Personal Model, Elephant State,
Episode / Loop / Step Trail, Contextual Recall, and Background Learning. Raw
Steps preserve what happened; contextual recall retrieves support from Steps
and Facts; proactive questions expose gaps, conflicts, stale Pulse, and useful
adaptations; background learning decides when evidence should change durable
claims. There is no free-form memory-note table or foreground memory write
surface. Durable understanding is written through governed Personal Model
tools and remains inspectable, correctable, and source-provenanced.

This report describes the Elephant Agent architecture, four-lens Personal Model
schema, proactive curiosity loop, local semantic recall path, background reflect
pipeline, dashboard inspection surfaces, and evaluation scenarios for wake
recovery, correction, no-match recall, and memory quality over time.
