\section{Personal-Model-First Memory}
\label{sec:abstraction}

Elephant Agent treats the Personal Model as the primary object of personal AI
improvement. A reusable skill can make an agent more capable, but it does not
by itself say what Elephant Agent currently understands about a specific user. The
Personal Model supplies that owner.

\subsection{Why Skill-First Evolution Is Incomplete}

Skill-first systems optimize for reusable workflows. They can discover that a
tool chain often succeeds, package the path, and make it easier to invoke
later. This is useful, but it leaves the core personal questions unresolved:

\begin{itemize}[leftmargin=*]
  \item What does Elephant Agent currently understand about this user?
  \item Which source Episodes support that understanding?
  \item How can the user correct, forget, or dispute it?
  \item Which question would materially improve future help?
\end{itemize}

These are not skill-package questions. They are Understanding System questions.
Elephant Agent therefore treats skills as ordinary capabilities, while durable
personalization lives in facts, questions, and provenance.

\subsection{Personal Model Contents}

The Personal Model is made of active facts grouped by four lenses:

\begin{center}
\begin{tabularx}{0.98\linewidth}{@{}lX@{}}
\toprule
\textbf{Lens} & \textbf{Purpose} \\
\midrule
Identity & Durable attributes: name, values, decision style, boundaries, and stable self-descriptions. \\
World & The user's environment: people, relationships, projects, tools, domains, places, and stable context. \\
Pulse & Current state: active work, current pressure, recent constraints, priorities, and mood patterns. \\
Journey & Accumulated experience: lessons, failures, recovery patterns, past decisions, and long-running growth. \\
\bottomrule
\end{tabularx}
\end{center}

Every fact has a lens, topic, text, status, confidence, source, and
\texttt{source\_episode\_ids}. Only active facts enter the stable prompt.
Retired and disputed facts remain available for audit but must not shape
answers.

\subsection{Remember Less, Understand Deeper}

The Personal Model is not a complete profile. It is a compact set of claims that
matter for future help. The system should preserve information when it changes
future collaboration, recall, risk judgment, or continuity. It should not
convert every chat detail into durable truth.

This gives the product rule a technical form: remember less, understand deeper.
The durable unit is not a transcript sentence. The durable unit is a correctable
fact with source Episode provenance and an explicit lens/topic.

\subsection{Capabilities as Downstream Tools}

Skills remain useful as operator-controlled procedures. They can be installed,
viewed, enabled, disabled, and invoked when relevant. They are not the owner of
personal truth and are not ranked by a separate hidden affinity model.

The result is a simpler inversion: Elephant Agent does not ask only what the agent can
learn to do. It asks what it currently understands about the person, why that
understanding exists, and how the user can correct it.
