\section{Conclusion}
\label{sec:conclusion}

Elephant Agent implements Personal-Model-first memory for personal AI. The
system does not treat reusable skill acquisition or raw recall as the primary
improvement target. It treats the Personal Model as the center: the durable,
grounded model of the user across Identity, World, Pulse, and Journey.

The clean Understanding System gives that center a concrete path:

\begin{center}
\texttt{Personal Model -> Elephant State -> Step Trail -> Contextual Recall -> Background Learning}
\end{center}

The Personal Model stores active facts and open questions. Elephant State keeps
identity and one current context note. The Episode / Loop / Step trail records
atomic source material. Contextual Recall retrieves support without turning it
into prompt truth. Background Learning reads Steps directly, writes through the
same foreground tools, and persists learning results for inspection.

This structure changes how personal AI improves. Resuming is not transcript
replay, and learning is not only adding reusable skills. Instead, explicit
Personal Model updates are controlled, proactive curiosity operates at the
user's chosen effort level, inferred personalization matures through reflection,
and current-turn recall searches active claims with no-match as a safe outcome.

The practical outcome is a personal AI that can explain what happened, continue
the right line of work, update long-term understanding through governed
provenance, repair itself after correction, and use capabilities while keeping
durable understanding inspectable. Skills still matter. In Elephant Agent, they remain
downstream operator-managed tools around an evolving Personal Model.
