\section{Related Systems and Comparison Axes}
\label{sec:related-systems}

This report does not add unverified citations. A citation-verified related
systems pass should be added later through programmatic reference checks rather
than memory-based BibTeX entries. The comparison axes are still clear.

\begin{itemize}[leftmargin=*]
  \item \textbf{Stateless chat systems.} These optimize the current prompt but
  do not preserve durable collaboration state.

  \item \textbf{Long-context systems.} These extend the input window but do not
  by themselves define what should become State or Personal Model truth.

  \item \textbf{Memory-augmented assistants.} These retrieve prior snippets,
  but often leave explicit capture, inferred personalization, maturity,
  correction, and scope boundaries underspecified.

  \item \textbf{Workflow agents.} These execute tool-heavy tasks, but often
  treat persistence as logs and checkpoints rather than as a layered personal
  model.

  \item \textbf{Companion systems.} These move closer to Elephant Agent when they model
  the user, relationship, and elephant identity boundaries, but they often blur
  Episode, State, Personal Model, and procedural learning.

  \item \textbf{Skill and automation systems.} These make repeated workflows
  reusable, but they rarely define the boundary between durable personal
  understanding and executable skill artifacts.

  \item \textbf{Skill-first self-evolving agents.} These optimize for acquiring
  reusable skills, tools, or workflows. Elephant Agent treats that as useful capability
  work, but not as the center of personal AI. The center is the Personal Model:
  active facts, source provenance, and useful questions.
\end{itemize}

Elephant Agent is therefore best compared on ownership boundaries. Atomic events belong
in Steps. Execution belongs in Loops. Wake windows belong in Episodes.
Workline continuity belongs in State. Long-term personalization belongs in the
Personal Model. Background reflect turns Step source material into fact changes
and question candidates without collapsing learned truth into a skill package.
